\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}

\signature{Corresponding author name\\Corresponding author affiliation\\Corresponding author email}
\address{Corresponding author address}
\date{May 16, 2026}

\begin{document}

\begin{letter}{Editors-in-Chief\\SIAM Journal on Financial Mathematics}

\opening{Dear Editors,}

Please consider the manuscript ``Structural Ambiguity Amplification in
Near-Critical Hawkes Market Making'' for publication as a full-length research
article in \emph{SIAM Journal on Financial Mathematics}.

The manuscript studies structural uncertainty in Hawkes-driven market making.
Its central contribution is a reproducible mathematical and computational
package showing how ambiguity in the branching matrix is amplified by the
Hawkes resolvent near criticality. The paper distinguishes relative
critical-slack ambiguity from absolute branching-matrix ambiguity, verifies a
finite-state robust Bellman layer and a multitype Hawkes HJBI/control bridge
for the stated exponential model class, and evaluates the resulting policy
diagnostics on simulated Hawkes paths, public LOBSTER samples, Binance
aggregate-trade events, and crypto level-2 depth data.

The accompanying reproducibility package contains source code, tests,
experiment scripts, public-data diagnostics, SOTA comparison tables, reference
audits, and a strict submission validator. The current validator checks
artifact presence, TeX logs, citation integrity, SOTA metadata, displayed-depth
queue replay diagnostics, M-series benchmark controls, and proof/scope audit
anchors.

To the best of our knowledge, this manuscript is original, is not under review
elsewhere, and has not been published previously.
% TODO before submission: disclose any conference, preprint, or repository
% history here if applicable.

Artificial intelligence disclosure draft: portions of the manuscript drafting,
code scaffolding, experiment automation, documentation, and validation scripts
were prepared with assistance from OpenAI Codex, a GPT-5 based coding
assistant, on May 16, 2026. The human authors are responsible for all
mathematical claims, proofs, citations, code, data processing, numerical
experiments, and final text included in the submitted manuscript.

Thank you for your consideration.

\closing{Sincerely,}

\end{letter}

\end{document}
